---
title: "Deliverable"
author: "Leah Bragg"
---

```{r, results='hide'}
knitr::opts_chunk$set(fig.width=5, fig.height=3)

suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(stringr)))
suppressWarnings(suppressMessages(library(tigris)))
```

## Data
```{r}
# read_excel("FoodAccessResearchAtlasData2019.xlsx",
#             "Food Access Research Atlas",
#             na = "NULL") |> filter(State == "Arkansas") |>
#    write_csv("arkansas-atlas2019.csv")
# 
# read_excel("FoodAccessResearchAtlasData2010.xlsx",
#            "Food Access Research Atlas",
#            na = "NULL") |> filter(State == "AR") |>
#   write_csv("arkansas-atlas2010.csv")

data2010 <- read_csv("arkansas-atlas2010.csv", 
                     show_col_types = FALSE) |>
  mutate(year = 2010, NAME = County)

data2019 <- read_csv("arkansas-atlas2019.csv", 
                     show_col_types = FALSE) |>
  mutate(year = 2019, NAME = word(County , 1  , -2))

```

## Statistical Test

H0: The mean number of people living 10 miles away from food (by county) has remained constant from 2010 to 2019.

Ha: The mean number of people living 10 miles away from food (by county) has shifted from 2010 to 2019.

```{r}
mutated_2010 <- data2010 |>
  mutate(num2010 = lapop10) |>
  select(NAME, num2010) |> group_by(NAME) |>
  summarize(num2010 = sum(num2010, na.rm = TRUE))

mutated_2019 <- data2019 |>
  mutate(num2019 = lapop10) |>
  select(NAME, num2019) |> group_by(NAME) |>
  summarize(num2019 = sum(num2019, na.rm = TRUE))

num2010_19 <- left_join(mutated_2010,
                                mutated_2019,
                                by = "NAME")
```

```{r}
t.test(num2010_19$num2010,
       num2010_19$num2019, paired = TRUE)
```

p-value = 0.1744, does not reject H0 at 0.05 level

```{r}
t.test(sqrt(num2010_19$num2010),
       sqrt(num2010_19$num2019), paired = TRUE)
```

p-value = 0.09216, does not reject H0 at 0.05 level

```{r}
wilcox.test(num2010_19$num2010,
       num2010_19$num2019, paired = TRUE)
```

p-value = 0.2296, does not reject H0 at 0.05 level


### lapop10 histograms/density plots
```{r}
for_hist2010 <- data2010 |>
  select(NAME, lapop10, year) |> group_by(NAME) |>
  summarize(lapop10 = sum(lapop10, na.rm = TRUE), 
            year = mean(year), .groups = "drop")

for_hist2019 <- data2019 |>
  select(NAME, lapop10, year) |> group_by(NAME) |>
  summarize(lapop10 = sum(lapop10, na.rm = TRUE), 
            year = mean(year), .groups = "drop")

hist_2010_19 <- bind_rows(for_hist2010, for_hist2019)

hist_2010_19 |> mutate(year = factor(year)) |>
  ggplot(aes(x = lapop10, color = year)) +
  geom_density(linewidth = 1) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  xlab("Population living 10 miles from food")

hist_2010_19 |> mutate(year = factor(year)) |>
  ggplot(aes(x = sqrt(lapop10), color = year)) +
  geom_density(linewidth = 1) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  xlab("Square root of Population living 10 miles from food")

num2010_19 |> ggplot(aes(x = num2019 - num2010)) +
  geom_histogram(fill = "deepskyblue4") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Difference from 2010 to 2019")
```
### lalowi10 histograms/density plots
```{r}
for_hist2010 <- data2010 |>
  select(NAME, lalowi10, year) |> group_by(NAME) |>
  summarize(lalowi10 = sum(lalowi10, na.rm = TRUE), 
            year = mean(year), .groups = "drop")

for_hist2019 <- data2019 |>
  select(NAME, lalowi10, year) |> group_by(NAME) |>
  summarize(lalowi10 = sum(lalowi10, na.rm = TRUE), 
            year = mean(year), .groups = "drop")

hist_2010_19 <- bind_rows(for_hist2010, for_hist2019)

hist_2010_19 |> mutate(year = factor(year)) |>
  ggplot(aes(x = lalowi10, color = year)) +
  geom_density(linewidth = 1) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  xlab("Low income population living 10 miles from food")

hist_2010_19 |> mutate(year = factor(year)) |>
  ggplot(aes(x = sqrt(lalowi10), color = year)) +
  geom_density(linewidth = 1) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  xlab("square root of low income population living 10 miles from food")

```

## Maps

```{r, results='hide'}
ar_map2010 <- tigris::counties(state = "Arkansas", 
                               year = 2010) |>
  mutate(NAME = NAME10, year = 2010)
ar_map2019 <- tigris::counties(state = "Arkansas", 
                               year = 2019) |>
  mutate(year = 2019)
ar_map2010_19 <- bind_rows(ar_map2010, ar_map2019)
```

### lapop10 maps
```{r}
dataForMap_2010 <- data2010 |> select(NAME, 
                                 lapop10)|>
  group_by(NAME) |> 
  summarize(value = sum(lapop10, na.rm = TRUE)) |>
  mutate(year = 2010)

dataForMap_2019 <- data2019 |> select(NAME, 
                                 lapop10, year) |>
  group_by(NAME) |> 
  summarize(value = sum(lapop10, na.rm = TRUE)) |>
  mutate(year = 2019)

dataForMap <- bind_rows(dataForMap_2010,
                        dataForMap_2019)

d <- left_join(ar_map2010_19, 
               dataForMap, 
               by = c("NAME" = "NAME", "year" = "year"))
d |> 
  ggplot(aes(fill = value)) +
  facet_wrap(vars(year)) +
  geom_sf() +
  theme_void() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) 

```

### lalowi10 maps
```{r}
dataForMap_2010 <- data2010 |> select(NAME, 
                                 lalowi10)|>
  group_by(NAME) |> 
  summarize(value = sum(lalowi10, na.rm = TRUE)) |>
  mutate(year = 2010)

dataForMap_2019 <- data2019 |> select(NAME, 
                                 lalowi10, year) |>
  group_by(NAME) |> 
  summarize(value = sum(lalowi10, na.rm = TRUE)) |>
  mutate(year = 2019)

dataForMap <- bind_rows(dataForMap_2010,
                        dataForMap_2019)

d <- left_join(ar_map2010_19, 
               dataForMap, 
               by = c("NAME" = "NAME", "year" = "year"))
d |> 
  ggplot(aes(fill = value)) +
  facet_wrap(vars(year)) +
  geom_sf() +
  theme_void() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) 

```

### lalowi10 maps
```{r}
dataForMap_2010 <- data2010 |> select(NAME, 
                                 LowIncomeTracts)|>
  group_by(NAME) |> 
  summarize(value = sum(LowIncomeTracts, na.rm = TRUE)) |>
  mutate(year = 2010)

dataForMap_2019 <- data2019 |> select(NAME, 
                                 LowIncomeTracts, year) |>
  group_by(NAME) |> 
  summarize(value = sum(LowIncomeTracts, na.rm = TRUE)) |>
  mutate(year = 2019)

dataForMap <- bind_rows(dataForMap_2010,
                        dataForMap_2019)

d <- left_join(ar_map2010_19, 
               dataForMap, 
               by = c("NAME" = "NAME", "year" = "year"))
d |> 
  ggplot(aes(fill = value)) +
  facet_wrap(vars(year)) +
  geom_sf() +
  theme_void() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) 

```

### 2019 TractLOWI
```{r}
dataForMap_2019 <- data2019 |> select(NAME, 
                                 TractLOWI, year) |>
  group_by(NAME) |> 
  summarize(value = sum(TractLOWI, na.rm = TRUE)) |>
  mutate(year = 2019)



d <- left_join(ar_map2019, 
               dataForMap_2019, 
               by = c("NAME" = "NAME", "year" = "year"))
d |> 
  ggplot(aes(fill = value)) +
  facet_wrap(vars(year)) +
  geom_sf() +
  theme_void() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) 

```

## Top 10 worst lapop10
```{r}
mutated_2010 |> slice_max(num2010, n = 10)
mutated_2019 |> slice_max(num2019, n = 10)
```















